---
title: "Set up RF"
output: html_document
editor_options: 
  chunk_output_type: console
---

# load libraries

```{r}
library(pacman)
p_load(lubridate, tidyverse, dplyr, timetools, data.table, chron)
op <- options(digits.secs = 3)
```

# calculate predictor variables to go into RF

```{r}
# csv file with raw acc data is converted to DDMT format and raw values are converted to g by dividing all axes with 16384
# this file is then loaded into DDMT, where I do an offset correction and set the device orientation
# corrected ACC values are then saved as a text file
# this text file is then manually converted to csv (also format ACC columns as text!) (can I load this file directly?)
# then load this file with corrected acc values in g
# with this I can calculate predictor variables


acc <- fread("E:/Myotis_vivesi/DDMT/Mviv18_06_ACC_Call_split1.txt")

#### only for 17_49 with two splits from DDMT
acc2 <- fread("E:/Myotis_vivesi/DDMT/Mviv17_49_ACC_Call_split2.txt")

names(acc1)
names(acc1)[1] <- "Event"

names(acc2)
names(acc2)[1] <- "Event"

acc <- rbind(acc1, acc2)
acc <- acc[order(acc$Event), ]
acc <- acc %>% distinct(Event, .keep_all = TRUE)
####


names(acc)
acc <- acc[,1:6]
names(acc)[1] <- "Event"
names(acc)[2] <- "ACCX"
names(acc)[3] <- "ACCY"
names(acc)[4] <- "ACCZ"
names(acc)[6] <- "Time"

acc$datetime <- paste(acc$Date, acc$Time)

acc <- acc[, -c(5:6)]


# calculate sampling rate
acc$datetime <- dmy_hms(acc$datetime)
acc$datetime <- acc$datetime - 7*3600
acc <- na.omit(acc)

options(digits.secs = 0)
AccHz <- median(table(acc$datetime))

# Hz states for the sampling frequency of acceleration (in our case 40Hz)
# Duration states for the timestep in seconds to be used in the running mean (in our case 3 seconds)
# Hz=40
# Duration=3
AccHz = Hz*Duration/2


# running mean 
Mov.Av. = function(obj){
  ma = obj
  for(i in AccHz:length(obj)){
    ma[i] = mean(obj[(i-AccHz):(i+AccHz)], na.rm = TRUE)
  }
  for(i in 1:AccHz){
    ma[i] = NA
  }
  for(i in (length(obj)-AccHz):length(obj)){
    ma[i] = NA
  }
  ma
}


# Static: running mean = 40 data points -> 1 sec
 acc$stX = Mov.Av.(acc$ACCX)
 acc$stY = Mov.Av.(acc$ACCY) 
 acc$stZ = Mov.Av.(acc$ACCZ)
 
# Dynamic
 acc$dX = acc$ACCX - acc$stX
 acc$dY = acc$ACCY - acc$stY
 acc$dZ = acc$ACCZ - acc$stZ
 
# VeDBA
 acc$VeDBA = sqrt(acc$dX^2 + acc$dY^2 + acc$dZ^2)
 
# VeDBAs
 # acc$VeDBAs = Mov.Av.(acc$VeDBA)

# VeSBA
 acc$VeSBA = sqrt(acc$stX^2 + acc$stY^2 + acc$stZ^2)
 
# Pitch
 acc$Pitch = asin(acc$stX)
 
# Roll
 acc$Roll = asin(acc$stY)
 
# PDBA
 #acc$PDBAX = abs(acc$dX)
 #acc$PDBAY = abs(acc$dY)
 #acc$PDBAZ = abs(acc$dZ)


options(digits.secs = 3)

write.csv(acc, "E:/Myotis_vivesi/18_06/Mviv18_06_acc_var.csv", row.names = FALSE)

detach("package:chron", unload = TRUE)

```

# add behaviour indexes: for each call find the matching acc and add an index

# add buzz idx

```{r}
# read in acc data
 acc <- read.csv("E:/Myotis_vivesi/18_06/Mviv18_06_acc_var.csv")

 acc$datetime <- ymd_hms(acc$datetime)

# fix offset file
buzz_times <- read.csv("E:/Myotis_vivesi/data/Audio/Mviv18/Mviv18_06/truebuzz/Mviv18_06_truebuzz.csv")
offset <- fread("E:/Myotis_vivesi/18_06/Mviv18_06_truebuzz_offset.csv")

names(offset)
names(offset)[1] <- "file"
names(offset)[2] <- "offset"
names(offset)[3] <- "loud_buzz_end"
names(offset)[4] <- "start_clip"
names(offset)[5] <- "segment"

offset$file <- buzz_times$file  
offset$segment <- buzz_times$segment
offset$start_clip <- ymd_hms(buzz_times$start_clip)

# replace Nas in loud buzz end with 0
# offset$loud_buzz_end <- if_else(is.na(offset$loud_buzz_end), 0, 1)

write.csv(offset, file = "E:/Myotis_vivesi/18_06/Mviv18_06_truebuzz_offset1.csv", row.names = FALSE)


# read in true_buzz_offset and calculate true_buzz_ends
true_buzz <- read.csv("E:/Myotis_vivesi/18_06/Mviv18_06_truebuzz_offset1.csv")

true_buzz$start_clip <- ymd_hms(true_buzz$start_clip)

#calculate end of buzz in each clip
true_buzz$buzz_end <- NA
true_buzz$buzz_end <- true_buzz$start_clip + true_buzz$segment*0.5 + true_buzz$offset

true_buzz$local_buzz_end <- NA
true_buzz$local_buzz_end <- true_buzz$buzz_end - 7*3600

true_buzz <- true_buzz[order(true_buzz$local_buzz_end),]

write.csv(true_buzz, "E:/Myotis_vivesi/18_06/Mviv18_06_true_buzz_end.csv", row.names = FALSE)


# read in true_buzz_end file
true_buzz_end <- read.csv("E:/Myotis_vivesi/18_06/Mviv18_06_true_buzz_end.csv")

# remove NAs
true_buzz_end <- na.omit(true_buzz_end)

true_buzz_end$local_buzz_end <- ymd_hms(true_buzz_end$local_buzz_end)

# check for duplicates, file length should match output
true_buzz_end$local_buzz_end %>% unique %>% length

true_buzz_end <- true_buzz_end[!duplicated(true_buzz_end$local_buzz_end), ]


acc$buzz_idx <- NA

for(i in 1:nrow(true_buzz_end)){
  buzz_end <- acc$datetime[which.min(abs(true_buzz_end$local_buzz_end[i] - acc$datetime))]
  start <- which.min(abs((buzz_end - 0.25) - acc$datetime))
  end <- which.min(abs((buzz_end + 0.25) - acc$datetime))
  acc$buzz_idx[start:end] <- i
}

```

# add comm idx

```{r}
# change 2017 to 2018 for 18_06
get_time <- function(file, year = 2018, ...){
    if(nchar(file) > 30){
      split <- strsplit(file, split = "_")
      month <- as.numeric(split[[1]][4])
      day <- as.numeric(split[[1]][5])
      hour <- as.numeric(split[[1]][6])
      min <- as.numeric(split[[1]][7])
      sec <- as.numeric(split[[1]][8])
      part <- as.numeric(substr(split[[1]][length(split[[1]])], 1, 
                                nchar(split[[1]][length(split[[1]])])-4))
        
      start <- ymd(paste0(year, "-", month, "-", day)) + hours(hour) + minutes(min) +
        seconds(sec) + seconds(part*0.5)
      return(as.character(start))  
    }
}


# function for 17_49
get_time <- function(file, year = 2017, ...){
    if(nchar(file) > 30){
      split <- strsplit(file, split = "_")
      month <- as.numeric(split[[1]][5])
      day <- as.numeric(split[[1]][6])
      hour <- as.numeric(split[[1]][7])
      min <- as.numeric(split[[1]][8])
      sec <- as.numeric(split[[1]][9])
      part <- as.numeric(substr(split[[1]][length(split[[1]])], 1, 
                                nchar(split[[1]][length(split[[1]])])-4))
        
      start <- ymd(paste0(year, "-", month, "-", day)) + hours(hour) + minutes(min) +
        seconds(sec) + seconds(part*0.5)
      return(as.character(start))  
    }
}


# get commuting times from audio files
comm_files <- list.files("E:/Myotis_vivesi/data/Audio/Mviv18/Mviv18_06/Commute", 
                         include.dirs = FALSE, pattern = "Mviv*")
comm_times <- data.frame(filename = comm_files)
file <- comm_files[1]

get_time(comm_files[1])

# set start and end times for each audio snippet
comm_times$start <- ymd_hms(sapply(X = comm_files, FUN = get_time)) - 7*3600
comm_times$end <- comm_times$start + 0.5

# remove NAs
comm_times <- na.omit(comm_times)

# check for duplicates, file length should match output
comm_times$start %>% unique %>% length

comm_times <- comm_times[!duplicated(comm_times$start), ]

comm_times <- comm_times[order(comm_times$start), ]


acc$comm_idx <- NA

for(i in 1:nrow(comm_times)){
  start <- which.min(abs(comm_times$start[i] - acc$datetime))
  end <- which.min(abs(comm_times$end[i] - acc$datetime))
  acc$comm_idx[start:end] <- i
}

```

# add roost idx

```{r}
# get roost times from audio files
roost_files <- list.files("E:/Myotis_vivesi/data/Audio/Mviv18/Mviv18_06/Roost", 
                         include.dirs = FALSE, pattern = "Mviv*")
roost_times <- data.frame(filename = roost_files)
file <- roost_files[1]

get_time(roost_files[1])

# set start and end times for each audio snippet
roost_times$start <- ymd_hms(sapply(X = roost_files, FUN = get_time)) - 7*3600
roost_times$end <- roost_times$start + 0.5


#### for 17_49
#1:184 Mviv17_49_021082_U17_06_23_08_43_48.346_363U-U17_06_22_08_43_45.498_363U.wav_23.jpg
#185:rest Mviv17_49_U17_06_22_08_43_50.494_363U-U17_06_22_08_43_54.782_363U.wav_000.jpg
roost_times$start <- NA
roost_times$start[1] <- get_time(roost_files[1])

roost_times$start[1:184] <- sapply(X = roost_files[1:184], FUN = get_time)
roost_times$start[185:499] <- sapply(X = roost_files[185:499], FUN = get_time1)
roost_times$start <- ymd_hms(roost_times$start)
roost_times$start <- roost_times$start - 7*3600
roost_times$end <- roost_times$start + 0.5
####


# remove NAs
roost_times <- na.omit(roost_times)

# check for duplicates, file length should match output
roost_times$start %>% unique %>% length

roost_times <- roost_times[!duplicated(roost_times$start), ]

roost_times <- roost_times[order(roost_times$start),]


acc$roost_idx <- NA

for(i in 1:nrow(roost_times)){
  start <- which.min(abs(roost_times$start[i] - acc$datetime))
  end <- which.min(abs(roost_times$end[i] - acc$datetime))
  acc$roost_idx[start:end] <- i
}

```

# add search idx

```{r}
# get search times from audio files
search_files <- list.files("E:/Myotis_vivesi/data/Audio/Mviv18/Mviv18_06/Search", 
                         include.dirs = FALSE, pattern = "Mviv*")
search_times <- data.frame(filename = search_files)
file <- search_files[1]

get_time(search_files[1])

# set start and end times for each audio snippet
search_times$start <- ymd_hms(sapply(X = search_files, FUN = get_time)) - 7*3600
search_times$end <- search_times$start + 0.5

# remove NAs
search_times <- na.omit(search_times)

# check for duplicates, file length should match output
search_times$start %>% unique %>% length

search_times <- search_times[!duplicated(search_times$start), ]

search_times <- search_times[order(search_times$start), ]


acc$search_idx <- NA

for(i in 1:nrow(search_times)){
  start <- which.min(abs(search_times$start[i] - acc$datetime))
  end <- which.min(abs(search_times$end[i] - acc$datetime))
  acc$search_idx[start:end] <- i
}


write.csv(acc, "E:/Myotis_vivesi/18_06/Mviv18_06_acc_var_idx.csv", row.names = FALSE)

```

# clean up indexes

```{r}
acc <- read.csv("E:/Myotis_vivesi/18_06/Mviv18_06_acc_var_idx.csv")

# buzz idx
buzz_count <- count(acc, buzz_idx)

table(buzz_count$n)

acc$buzz_idx_clean <- NA

# for 40 Hz ~ 20; for 50 Hz ~ 25

for (i in 1:nrow(buzz_count)){
  if(buzz_count$n[i] == 25) {
    acc$buzz_idx_clean[which(acc$buzz_idx == buzz_count$buzz_idx[i])] <- i
  }
}

for (i in 1:nrow(buzz_count)){
  if(buzz_count$n[i] == 26) {
    acc$buzz_idx_clean[which(acc$buzz_idx == buzz_count$buzz_idx[i])] <- i
    }
}

buzz_count_clean <- count(acc, buzz_idx_clean)

table(buzz_count_clean$n)


# comm_idx
comm_count <- count(acc, comm_idx)

table(comm_count$n)

acc$comm_idx_clean <- NA

for (i in 1:nrow(comm_count)){
  if(comm_count$n[i] == 25) {
    acc$comm_idx_clean[which(acc$comm_idx == comm_count$comm_idx[i])] <- i
  }
}

for (i in 1:nrow(comm_count)){
  if(comm_count$n[i] == 26) {
    acc$comm_idx_clean[which(acc$comm_idx == comm_count$comm_idx[i])] <- i
    }
}

comm_count_clean <- count(acc, comm_idx_clean)

table(comm_count_clean$n)


# search_idx
search_count <- count(acc, search_idx)

table(search_count$n)

acc$search_idx_clean <- NA

for (i in 1:nrow(search_count)){
  if(search_count$n[i] == 25) {
    acc$search_idx_clean[which(acc$search_idx == search_count$search_idx[i])] <- i
    }
}

for (i in 1:nrow(search_count)){
  if(search_count$n[i] == 26) {
    acc$search_idx_clean[which(acc$search_idx == search_count$search_idx[i])] <- i
    }
}

search_count_clean <- count(acc, search_idx_clean)

table(search_count_clean$n)


# roost_idx
roost_count <- count(acc, roost_idx)

table(roost_count$n)

acc$roost_idx_clean <- NA

for (i in 1:nrow(roost_count)){
  if(roost_count$n[i] == 25) {
    acc$roost_idx_clean[which(acc$roost_idx == roost_count$roost_idx[i])] <- i
  }
}

for (i in 1:nrow(roost_count)){
  if(roost_count$n[i] == 26) {
    acc$roost_idx_clean[which(acc$roost_idx == roost_count$roost_idx[i])] <- i
    }
}

roost_count_clean <- count(acc, roost_idx_clean)

table(roost_count_clean$n)


write.csv(acc, "E:/Myotis_vivesi/18_06/Mviv18_06_acc_var_idx_clean.csv", row.names = FALSE)

```

# create 0.5 s files by summarizing the acc data by indexes

# summarize buzz idx

```{r}
acc <- read.csv("E:/Myotis_vivesi/18_06/Mviv18_06_acc_var_idx_clean.csv")

buzz_acc <- acc %>% group_by(buzz_idx_clean) %>% 
  summarise(count = n(),
            mean.X = mean(ACCX), mean.Y = mean(ACCY), mean.Z = mean(ACCZ),
            max.X = max(ACCX), max.Y = max(ACCY), max.Z = max(ACCZ),
            min.X = min(ACCX), min.Y = min(ACCY), min.Z = min(ACCZ),
            range.X = diff(range(ACCX)), range.Y = diff(range(ACCY)), range.Z = diff(range(ACCZ)),
            mean.stX = mean(stX), mean.stY = mean(stY), mean.stZ = mean(stZ),
            max.stX = max(stX), max.stY = max(stY), max.stZ = max(stZ),
            min.stX = min(stX), min.stY = min(stY), min.stZ = min(stZ),
            range.stX = diff(range(stX)), range.stY = diff(range(stY)), range.stZ = diff(range(stZ)),
            mean.dX = mean(dX), mean.dY = mean(dY), mean.dZ = mean(dZ),
            max.dX = max(dX), max.dY = max(dY), max.dZ = max(dZ),
            min.dX = min(dX), min.dY = min(dY), min.dZ = min(dZ),
            range.dX = diff(range(dX)), range.dY = diff(range(dY)), range.dZ = diff(range(dZ)),
            mean.VeDBA = mean(VeDBA), mean.VeSBA = mean(VeSBA),
            max.VeDBA = max(VeDBA), max.VeSBA = max(VeSBA),
            min.VeDBA = min(VeDBA), min.VeSBA = min(VeSBA),
            range.VeDBA = diff(range(VeDBA)), range.VeSBA = diff(range(VeSBA)),
            mean.Pitch = mean(Pitch), max.Pitch = max(Pitch), min.Pitch = min(Pitch), range.Pitch = diff(range(Pitch)),
            mean.Roll = mean(Roll), max.Roll = max(Roll), min.Roll = min(Roll), range.Roll = diff(range(Roll)))

head(buzz_acc, 10)

# add behavior
buzz_acc$behav = "buzz"

table(buzz_acc$count)

names(buzz_acc)
buzz_acc <- buzz_acc[, -c(1,2)]

write.csv(buzz_acc, "E:/Myotis_vivesi/18_06/Mviv18_06_buzz_acc.csv", row.names = FALSE)

```

# summarize comm idx

```{r}
comm_acc <- acc %>% group_by(comm_idx_clean) %>% 
  summarise(count = n(),
            mean.X = mean(ACCX), mean.Y = mean(ACCY), mean.Z = mean(ACCZ),
            max.X = max(ACCX), max.Y = max(ACCY), max.Z = max(ACCZ),
            min.X = min(ACCX), min.Y = min(ACCY), min.Z = min(ACCZ),
            range.X = diff(range(ACCX)), range.Y = diff(range(ACCY)), range.Z = diff(range(ACCZ)),
            mean.stX = mean(stX), mean.stY = mean(stY), mean.stZ = mean(stZ),
            max.stX = max(stX), max.stY = max(stY), max.stZ = max(stZ),
            min.stX = min(stX), min.stY = min(stY), min.stZ = min(stZ),
            range.stX = diff(range(stX)), range.stY = diff(range(stY)), range.stZ = diff(range(stZ)),
            mean.dX = mean(dX), mean.dY = mean(dY), mean.dZ = mean(dZ),
            max.dX = max(dX), max.dY = max(dY), max.dZ = max(dZ),
            min.dX = min(dX), min.dY = min(dY), min.dZ = min(dZ),
            range.dX = diff(range(dX)), range.dY = diff(range(dY)), range.dZ = diff(range(dZ)),
            mean.VeDBA = mean(VeDBA), mean.VeSBA = mean(VeSBA),
            max.VeDBA = max(VeDBA), max.VeSBA = max(VeSBA),
            min.VeDBA = min(VeDBA), min.VeSBA = min(VeSBA),
            range.VeDBA = diff(range(VeDBA)), range.VeSBA = diff(range(VeSBA)),
            mean.Pitch = mean(Pitch), max.Pitch = max(Pitch), min.Pitch = min(Pitch), range.Pitch = diff(range(Pitch)),
            mean.Roll = mean(Roll), max.Roll = max(Roll), min.Roll = min(Roll), range.Roll = diff(range(Roll)))

head(comm_acc, 10)

# add behavior
comm_acc$behav = "comm" 

table(comm_acc$count)

names(comm_acc)
comm_acc <- comm_acc[, -c(1,2)]

write.csv(comm_acc, "E:/Myotis_vivesi/18_06/Mviv18_06_comm_acc.csv", row.names = FALSE)

```

# summarize roost idx

```{r}
roost_acc <- acc %>% group_by(roost_idx_clean) %>% 
  summarise(count = n(),
            mean.X = mean(ACCX), mean.Y = mean(ACCY), mean.Z = mean(ACCZ),
            max.X = max(ACCX), max.Y = max(ACCY), max.Z = max(ACCZ),
            min.X = min(ACCX), min.Y = min(ACCY), min.Z = min(ACCZ),
            range.X = diff(range(ACCX)), range.Y = diff(range(ACCY)), range.Z = diff(range(ACCZ)),
            mean.stX = mean(stX), mean.stY = mean(stY), mean.stZ = mean(stZ),
            max.stX = max(stX), max.stY = max(stY), max.stZ = max(stZ),
            min.stX = min(stX), min.stY = min(stY), min.stZ = min(stZ),
            range.stX = diff(range(stX)), range.stY = diff(range(stY)), range.stZ = diff(range(stZ)),
            mean.dX = mean(dX), mean.dY = mean(dY), mean.dZ = mean(dZ),
            max.dX = max(dX), max.dY = max(dY), max.dZ = max(dZ),
            min.dX = min(dX), min.dY = min(dY), min.dZ = min(dZ),
            range.dX = diff(range(dX)), range.dY = diff(range(dY)), range.dZ = diff(range(dZ)),
            mean.VeDBA = mean(VeDBA), mean.VeSBA = mean(VeSBA),
            max.VeDBA = max(VeDBA), max.VeSBA = max(VeSBA),
            min.VeDBA = min(VeDBA), min.VeSBA = min(VeSBA),
            range.VeDBA = diff(range(VeDBA)), range.VeSBA = diff(range(VeSBA)),
            mean.Pitch = mean(Pitch), max.Pitch = max(Pitch), min.Pitch = min(Pitch), range.Pitch = diff(range(Pitch)),
            mean.Roll = mean(Roll), max.Roll = max(Roll), min.Roll = min(Roll), range.Roll = diff(range(Roll)))

head(roost_acc, 10)

# add behavior
roost_acc$behav = "roost" 

table(roost_acc$count)

names(roost_acc)
roost_acc <- roost_acc[, -c(1,2)]

write.csv(roost_acc, "E:/Myotis_vivesi/18_06/Mviv18_06_roost_acc.csv", row.names = FALSE)

```

# summarize search idx

```{r}
search_acc <- acc %>% group_by(search_idx_clean) %>% 
  summarise(count = n(),
            mean.X = mean(ACCX), mean.Y = mean(ACCY), mean.Z = mean(ACCZ),
            max.X = max(ACCX), max.Y = max(ACCY), max.Z = max(ACCZ),
            min.X = min(ACCX), min.Y = min(ACCY), min.Z = min(ACCZ),
            range.X = diff(range(ACCX)), range.Y = diff(range(ACCY)), range.Z = diff(range(ACCZ)),
            mean.stX = mean(stX), mean.stY = mean(stY), mean.stZ = mean(stZ),
            max.stX = max(stX), max.stY = max(stY), max.stZ = max(stZ),
            min.stX = min(stX), min.stY = min(stY), min.stZ = min(stZ),
            range.stX = diff(range(stX)), range.stY = diff(range(stY)), range.stZ = diff(range(stZ)),
            mean.dX = mean(dX), mean.dY = mean(dY), mean.dZ = mean(dZ),
            max.dX = max(dX), max.dY = max(dY), max.dZ = max(dZ),
            min.dX = min(dX), min.dY = min(dY), min.dZ = min(dZ),
            range.dX = diff(range(dX)), range.dY = diff(range(dY)), range.dZ = diff(range(dZ)),
            mean.VeDBA = mean(VeDBA), mean.VeSBA = mean(VeSBA),
            max.VeDBA = max(VeDBA), max.VeSBA = max(VeSBA),
            min.VeDBA = min(VeDBA), min.VeSBA = min(VeSBA),
            range.VeDBA = diff(range(VeDBA)), range.VeSBA = diff(range(VeSBA)),
            mean.Pitch = mean(Pitch), max.Pitch = max(Pitch), min.Pitch = min(Pitch), range.Pitch = diff(range(Pitch)),
            mean.Roll = mean(Roll), max.Roll = max(Roll), min.Roll = min(Roll), range.Roll = diff(range(Roll)))

head(search_acc, 10)

# add behavior
search_acc$behav = "search" 

table(search_acc$count)

names(search_acc)
search_acc <- search_acc[, -c(1,2)]

write.csv(search_acc, "E:/Myotis_vivesi/18_06/Mviv18_06_search_acc.csv", row.names = FALSE)

```

# put together behaviour files

```{r}

acc_behav <- rbind(buzz_acc, roost_acc, search_acc, comm_acc)
names(acc_behav)

write.csv(acc_behav, "E:/Myotis_vivesi/18_06/Mviv18_06_acc_behav.csv", row.names = FALSE)

```